The Concurrent Calculi Formalisation Benchmark

### 2024-02-21 short-paper/short-paper.tex
shortening the paragraph about Celf

Similarly to the authors of the POPLMark challenges, we seek to answer several questions:
\begin{enumerate}[label=\textbf{(Q\arabic*)},leftmargin=10mm]
\item\label{item:rq1} What is the current state of the art in tools for mechanising the meta-theory of process calculi?
\item\label{item:rq2} Which techniques and best practices can be recommended when starting new mechanisation projects involving process calculi?
\item\label{item:rq3} Which improvements are needed to make mechanisation tools more user-friendly with regards to the issues faced when mechanising process calculi?
\end{enumerate}

@@

In mechanised meta-theory, addressing linearity means choosing an appropriate
representation of the linear typing context.  While a linear context is perhaps
best seen as a multi-set, most proof assistants have better support for list
processing.  The latter representation is intuitive, but may require
establishing a large number of technical lemmata that are orthogonal to the
problem under study (in our case, proving type safety for
session types).  In this space, several designs are possible: one can label
occurrences of resources to constrain their usage (\eg~\cite{CicconeP20}); some
authors instead impose a multiset structure over lists
(\eg~\cite{Danielsson12,ChaudhuriLR19}). Alternatively, contexts can be
implemented as finite maps (as in \cite{Castro2020}), whose operations are
sensitive to a linear discipline. In all these cases, the effort required to
develop the infrastructure is significant.
%
One alternative strategy is to bypass the problem of context splitting
by adopting familiar ideas from algorithmic linear type checking. This
approach, referred to as ``typing with leftovers'', is exemplified
in~\cite{DBLP:conf/forte/ZalakainD21}.  Whatever the choice,
list-based encodings can be refined to be intrinsically-typed, if the
proof assistant supports dependent types
(see~\cite{Thiemann2019,CicconeP20,RouvoetPKV20}).
%
A radically different approach is to adopt a \emph{substructural}
meta-logical framework: the framework itself handles resource
distribution, and users need only map their linear operations to the
ones offered by the framework.  The linear function space will then
handle all context operations implicitly, including splitting and
substitution.  The only such framework is
\emph{Celf}~\cite{Schack-Nielsen:IJCAR08} (see the encoding of session
types in~\cite{Bock2016}); unfortunately, \emph{Celf} does not yet
fully support the verification of meta-theoretic properties.  A
compromise is the so called \emph{two-level} approach, \ie encoding a
substructural specification logic in a mainstream proof assistant and
then using that logic to state and prove linear properties (for a
recent example, see~\cite{Felty:MSCS21}).

>>>>

Similarly to the authors of the POPLMark challenges, we seek to answer several questions:
\begin{enumerate}[label=\textbf{(Q\arabic*)},leftmargin=10mm]
\item\label{item:rq1} What is the current state of the art in the mechanisation of  the meta-theory of process calculi?
\item\label{item:rq2} Which techniques and best practices can be recommended when starting new mechanisation projects involving process calculi?
\item\label{item:rq3} Which improvements are needed to make mechanisation tools more user-friendly with regards to the issues faced when mechanising process calculi?
\end{enumerate}

@@

In mechanised meta-theory, addressing linearity means choosing an appropriate
representation of the linear typing context.  While a linear context is perhaps
best seen as a multiset, most proof assistants have better support for list
processing.  The latter representation is intuitive, but may require
establishing a large number of technical lemmata that are orthogonal to the
problem under study (in our case, proving type safety for
session types).  In this space, several designs are possible: one can label
occurrences of resources to constrain their usage (\eg~\cite{CicconeP20}); some
authors instead impose a multiset structure over lists
(\eg~\cite{Danielsson12,ChaudhuriLR19}). Alternatively, contexts can be
implemented as finite maps (as in \cite{Castro2020}), whose operations are
sensitive to a linear discipline. In all these cases, the effort required to
develop the infrastructure is significant.
%
One alternative strategy is to bypass the problem of context splitting
by adopting familiar ideas from algorithmic linear type checking. This
approach, referred to as ``typing with leftovers'', is exemplified
in~\cite{DBLP:conf/forte/ZalakainD21}.  Whatever the choice,
list-based encodings can be refined to be intrinsically-typed, if the
proof assistant supports dependent types
(see~\cite{Thiemann2019,CicconeP20,RouvoetPKV20}).
%
A radically different approach is to adopt a \emph{substructural}
meta-logical framework, which handles resource
distribution implicitly, including splitting and
substitution: users need only map their linear operations to the
ones offered by the framework.  The only such framework is
\emph{Celf}~\cite{Schack-Nielsen:IJCAR08} (see the encoding of session
types in~\cite{Bock2016}); unfortunately, \emph{Celf} does not yet
fully support the verification of meta-theoretic properties.  A
compromise is the so called \emph{two-level} approach, \ie encoding a
substructural specification logic in a mainstream proof assistant and
then using that logic to state and prove linear properties (for a
recent example, see~\cite{Felty:MSCS21}).
